Agenda diversity and coverage bias: A quantitative approach to the agenda-setting theory

### 2018-07-13 agenda_diversity.tex
Version_con_minimas_correcciones_submiteada

\par To understand the ecosystem of information flow and opinion formation is one of the goals in these days. In this framework, many ingredients can be part of different mechanism of social influence. A major role in this ecosystem is played by the Mass Media outlets, which used to be the source of information of many of us.  Informed people tends to interact with each other, either via personal interactions or through social networks. In this scenario, becomes essential to understand the role of Mass Media Influence in a given social group.
\par  This question has been raised years ago. In the famous study performed in Chapel Hill during the US presidential elections in 1968 \cite{mccombs1972agenda}, Maxwell McCombs and Donald Shaw found that those aspects of public affairs that are prominent in the news become prominent among the public.
This study is considered the founding of the agenda-setting theory, which focus in the influence of mass media in public opinion.  
From \cite{mccombs2014agenda}, \textit{``The media agenda is the pattern of news coverage over a period of days, weeks (...) for a set of issues or other topic. In other words, the media agenda is a systematic compilation of the issues or topics presented to the public that identifies the degree of emphasis on these topics.''}. 
\par In its basic stage, the theory focus on the comparison between the topics coverage by the media and the public agenda, i.e. the topics that the public consider as priority.
It looks for answering the question if the media is able to set the public agenda, which would transform the media as an important actor in the formation of public opinion. 
This stage is sometimes called the ``first level agenda-setting''.  The very often quoted phrase of Bernard C. Cohen \textit{``The press may not be successful much of the time in telling people what to think, but it is stunningly successful in telling its readers what to think about.''} illustrates its object of study. With the emergency of the Internet, the end of agenda-setting were predicted due to the audience fragmentation onto multiple sources, which would virtually lead to a highly individualized agenda. However, it is based on two assumptions that are not necessarily true: the public spreads its attention in an homogeneous way across the multiple sources, and that the agendas of that sources are different \cite{mccombs2005look}.

@@

\par Since the irruption of internet in our lives and the access to big data, there were many attempts to analyze from a quantitative point of view the theory of Agenda Setting based on  the original work of McCombs \cite{mccombs1972agenda}. Many of them follows what we called a "static approach", i.e., by summarizing certain issues present in news media and looking for effects in the audience \cite{brians1996campaign, gerber2009does, coleman2007young}.
For instance, in Argentina, \cite{mitchelstein2016brecha} examines the difference between public and journalists preferences. 
Several works by E. Zunino, like \cite{zunino2010cobertura} and \cite{koziner2013cobertura}, studied the coverage given by the main newspapers of particular events which have happened during the former administration of Argentinian president \emph{Cristina Fernandez de Kirchner} (from 2007 to 2015), where the government started to confront with news organizations that were critical to its management as they were an opposition party \cite{mitchelstein2017information}

@@

\subsection{The Public and Twitter Agendas}

\par In order to have a proxy about how the Media Agenda is reaching the public, we look for the same topics in Google Trends and Twitter in the same period of time. We take advantage of topics' keywords on the one hand by making queries to the \emph{Google Trends} tool and getting the relative weight of \emph{Google} searches that people made about the identified topics, and in the other hand, by making queries to the advance search tool in the social media \emph{Twitter} in order to get the relative amount of related tweets .

@@

\par In figure \ref{fig:all_agenda} we show the time evolution of the topics with a bump chart of the Agendas. The bump chart provides  a very useful visualization tool for displaying the relative weight of the topics at the same time that their ranking. In figure \ref{fig:all_agenda} we also highlight some important events related to the dynamics of the topics. \footnote{Due to different characteristics in the search tool of Twitter, we adapted the queries employed here but preserving at least at we can the most important keywords. The queries employed by topic were: \\
Elections: elecciones + cambiemos + kirchner + massa + randazzo \\
Missing person: maldonado + otranto + gendarmeria + desaparicion \\
Former Planning minister: vido + desafuero + minnicelli + baratta \\
Current President: macri + cgt + laboral +  triaca \\
Social leader:  sala + cidh + tupac + amaru + pullen + llermanos + morales \\
Prosecutor’s death: nisman + amia + memorandum + timerman +  bonadio \\
Former Vice-President:  boudou + ciccone +  lijo + vandenbroele + carmona
}

@@

\par A lower value in the agenda's diversity is due to the fact that the most important topic attracts practically all the attention of the public and the media.
In the radar plots included in figure \ref{fig:shannon_entropy_agendas} we can see that two of these outliers (\textbf{a} and \textbf{d}) belong to the topic \emph{Elections} and are related with the primary and general legislative elections that took place in August 13th and October 22th. 
In all the agendas these points where detected as outliers except point (d) in Twitter Agenda. Why is that? The radar plot of the Twitter agenda for this day displays an association between the topic \emph{Elections} and the \emph{Current President} which could be decreasing the important of this topic.
Discussions in Twitter about elections appear also in point (c), when the other agendas seems to be more diverse. 
On the other hand, we focused in point (b), despite not being detected as a outlier, because it belong to the topic \emph{Missing person} and this date correspond to the rally that took place one month after the disappearance of Santiago Maldonado (see section \ref{sec:Context}). It is important to notice that the Shannon's Entropy of GT Agenda displays a minimum (collapsing agenda) not shown neither in the Media nor in the Twitter Agendas.
We emphasize the discussion about this topic because we can appreciate  interesting facts that appear along the analysis.

@@

\par On the other hand,  points \textbf{(a)} (we take this point due to be a local maximum despite not being an outlier) and \textbf{(b)} seem to show an interest of the public in the topic \emph{Missing person} which it is not reflected in the Media.  In figure \ref{fig:all_agenda} we can see that this topic becomes the most important in public's interest (both \textbf{GT} and \textbf{Tw}) days before that  it happen in the Media Agenda. It can be associates  with a social networks (like Facebook and Twitter) campaign  in favor  of the appearance of Santiago Maldonado ("The missing person") that took place on  August 26th. It was very important and was initially underestimated by the main Media outlets in Argentina (see section \ref{sec:Context}).

@@

\par Finally, in figure \ref{fig:topics_temporal_profiles} we also show word-clouds of topic's keywords but separating those who are more frequently mentioned by each newspaper, and filtering the words that are common to all and basically define the factual details of the topic.
Although most of the words are not relevant enough, some of them are quite interesting, as for instance the word \emph{represi\'on} (repression) when \emph{P\'agina 12} talks about the topic \emph{Missing Person} and the word \emph{Cristina} (Fernandez de Kirchner, former president) which is employed by all newspapers except \emph{P\'agina 12} when they  talk about the topic \emph{Former Planning minister} (see section \ref{sec:Context}).

@@

\subsubsection{\emph{Missing Person}}
\par \emph{Santiago Maldonado} vanished on August 1st after a minor clash between the Gendarmerie (Border Guards) and a group of Mapuches, which recognize as themselves the original population of an area in the Patagonia.
Since that event, the \emph{Mauricio Macri}'s administration was accused by several people as the responsible of a \textbf{forced disappearance}. 
\par A very massive campaign in social media took place on August 26th, under the motto "Where is Santiago Maldonado?", followed by two massive protest marches to the \emph{Plaza de Mayo} that took place on September 1st and October 1st, which the first one had a great repercussion due to several incidents that took place during the march.
\par The body of \emph{Santiago Maldonado} was found on 17th October in the \emph{Chubut} river, near the place where he was seen the last time, and the autopsy report told that \emph{Santiago Maldonado} had died from ‘asphyxia after being submerged,’ with no injuries on his body. 
However the responsibility of the current administration is still being discussed.

@@

\par \emph{Julio de Vido} was the Planning minister during the administration of \emph{Nestor Kirchner} and \emph{Cristina Fernandez de Kirchner} (2003-2015). In 2015, he was elected to integrate the Chamber of Deputies, which finally voted to strip \emph{De Vido} of his congressional immunity over corruption allegations and was immediatly jailed on October 27th.

@@

\par \emph{Milagro Sala} is an indigenous leader’s. 
She has been incarcerated under pre-trial detention ever since she was first detained in January 2016. She faces allegations of embezzlement related to government funding for housing projects managed by T\'upac Amaru, her social organization.
Sala accused the government of "violating her human rights", and several people think that she is a political prisoner of the \emph{Mauricio Macri} administration.

>>>>

\par To understand the ecosystem of information flow and opinion formation is one of the goals in these days. In this framework, many ingredients can be part of different mechanism of social influence. A major role in this ecosystem is played by the Mass Media outlets, which used to be the source of information of many of us.  Informed people tends to interact with each other, either via personal interactions or through social networks. In this scenario, becomes essential to understand the role of Mass Media Influence in a given social group.
\par  This question has been raised years ago. In the famous study performed in Chapel Hill during the US presidential elections in 1968 \cite{mccombs1972agenda}, Maxwell McCombs and Donald Shaw found that those aspects of public affairs that are prominent in the news become prominent among the public.
This study is considered the founding of the agenda-setting theory, which focus in the influence of mass media in public opinion.  
From \cite{mccombs2014agenda}, \textit{``The media agenda is the pattern of news coverage over a period of days, weeks (...) for a set of issues or other topic. In other words, the media agenda is a systematic compilation of the issues or topics presented to the public that identifies the degree of emphasis on these topics''}. 
\par In its basic stage, the theory focus on the comparison between the topics coverage by the media and the public agenda, i.e. the topics that the public consider as priority.
It looks for answering the question if the media is able to set the public agenda, which would transform the media as an important actor in the formation of public opinion. 
This stage is sometimes called the ``first level agenda-setting''.  The very often quoted phrase of Bernard C. Cohen \textit{``The press may not be successful much of the time in telling people what to think, but it is stunningly successful in telling its readers what to think about.''} illustrates its object of study. With the emergency of the Internet, the end of agenda-setting were predicted due to the audience fragmentation onto multiple sources, which would virtually lead to a highly individualized agenda. However, it is based on two assumptions that are not necessarily true: the public spreads its attention in an homogeneous way across the multiple sources, and that the agendas of that sources are different \cite{mccombs2005look}.

@@

\par Since the irruption of internet in our lives and the access to big data, there were many attempts to analyze from a quantitative point of view the theory of Agenda Setting based on  the original work of McCombs \cite{mccombs1972agenda}. Many of them follows what we called a ``static approach", i.e. by summarizing certain issues present in news media and looking for effects in the audience \cite{brians1996campaign, gerber2009does, coleman2007young}.
For instance, in Argentina, \cite{mitchelstein2016brecha} examines the difference between public and journalists preferences. 
Several works by E. Zunino, like \cite{zunino2010cobertura} and \cite{koziner2013cobertura}, studied the coverage given by the main newspapers of particular events which have happened during the former administration of Argentinian president \emph{Cristina Fern\'andez de Kirchner} (from 2007 to 2015), where the government started to confront with news organizations that were critical to its management as they were an opposition party \cite{mitchelstein2017information}.

@@

\subsection{The Google and Twitter Agendas}

\par In order to have a proxy about how the Media Agenda is reaching the public, we look for the same topics in \emph{Google} and \emph{Twitter} in the same period of time. We take advantage of topics' keywords on the one hand by making queries to the \emph{Google Trends} tool and getting the relative weight of Google searches that people made about the identified topics, and in the other hand, by making queries to the advance search tool in the social media Twitter in order to get the relative amount of related tweets.

@@

\par In figure \ref{fig:all_agenda} we show the time evolution of the topics with a bump chart of the Agendas. The bump chart provides  a very useful visualization tool for displaying the relative weight of the topics at the same time that their ranking. In figure \ref{fig:all_agenda} we also highlight some important events related to the dynamics of the topics. \footnote{Due to different characteristics in the search tool of Twitter, we adapted the queries employed here but preserving at least at we can the most important keywords. The queries employed by topic were: \\
Elections: elecciones + cambiemos + kirchner + massa + randazzo \\
Missing person: maldonado + otranto + gendarmer\'ia + desaparici\'on \\
Former Planning minister: vido + desafuero + minnicelli + baratta \\
Current President: macri + cgt + laboral +  triaca \\
Social leader:  sala + cidh + tupac + amaru + pullen + llermanos + morales \\
Prosecutor’s death: nisman + amia + memorandum + timerman +  bonadio \\
Former Vice-President:  boudou + ciccone +  lijo + vandenbroele + carmona
}

@@

\par A lower value in the agenda's diversity is due to the fact that the most important topic attracts practically all the attention of the public and the media.
In the radar plots included in figure \ref{fig:shannon_entropy_agendas} we can see that two of these outliers (\textbf{a} and \textbf{d}) belong to the topic \emph{Elections} and are related with the primary and general legislative elections that took place in August 13th and October 22th. 
In all the agendas these points where detected as outliers except point (d) in Twitter Agenda. Why is that? The radar plot of the Twitter agenda for this day displays an association between the topic \emph{Elections} and the \emph{Current President} which could be decreasing the important of this topic.
Discussions in Twitter about elections appear also in point (c), when the other agendas seems to be more diverse. 
On the other hand, we focused in point (b), despite not being detected as a outlier, because it belong to the topic \emph{Missing person} and this date correspond to the rally that took place one month after the disappearance of \emph{Santiago Maldonado} (see section \ref{sec:Context}). It is important to notice that the Shannon's Entropy of GT Agenda displays a minimum (collapsing agenda) not shown neither in the Media nor in the Twitter Agendas.
We emphasize the discussion about this topic because we can appreciate  interesting facts that appear along the analysis.

@@

\par On the other hand,  points \textbf{(a)} (we take this point due to be a local maximum despite not being an outlier) and \textbf{(b)} seem to show an interest of the public in the topic \emph{Missing person} which it is not reflected in the Media.  In figure \ref{fig:all_agenda} we can see that this topic becomes the most important in public's interest (both \textbf{GT} and \textbf{Tw}) days before that  it happen in the Media Agenda. It can be associates  with a social networks (like Facebook and Twitter) campaign  in favor  of the appearance of \emph{Santiago Maldonado} (``The missing person") that took place on  August 26th. It was very important and was initially underestimated by the main Media outlets in Argentina (see section \ref{sec:Context}).

@@

\par Finally, in figure \ref{fig:topics_temporal_profiles} we also show word-clouds of topic's keywords but separating those who are more frequently mentioned by each newspaper, and filtering the words that are common to all and basically define the factual details of the topic.
Although most of the words are not relevant enough, some of them are quite interesting, as for instance the word \emph{represi\'on} (repression) when \emph{P\'agina 12} talks about the topic \emph{Missing Person} and the word \emph{Cristina} (Fern\'andez de Kirchner, former president) which is employed by all newspapers except \emph{P\'agina 12} when they  talk about the topic \emph{Former Planning minister} (see section \ref{sec:Context}).

@@

\subsubsection{\emph{Missing Person}}
\par \emph{Santiago Maldonado} vanished on August 1st after a minor clash between the Gendarmerie (Border Guards) and a group of Mapuches, which recognize as themselves the original population of an area in the Patagonia.
Since that event, the \emph{Mauricio Macri}'s administration was accused by several people as the responsible of a \textbf{forced disappearance}. 
\par A very massive campaign in social media took place on August 26th, under the motto ``Where is Santiago Maldonado?", followed by two massive protest marches to the \emph{Plaza de Mayo} that took place on September 1st and October 1st, which the first one had a great repercussion due to several incidents that took place during the march.
\par The body of \emph{Santiago Maldonado} was found on 17th October in the \emph{Chubut} river, near the place where he was seen the last time, and the autopsy report told that \emph{Santiago Maldonado} had died from ``asphyxia after being submerged", with no injuries on his body. 
However the responsibility of the current administration is still being discussed.

@@

\par \emph{Julio de Vido} was the Planning minister during the administration of \emph{N\'estor Kirchner} and \emph{Cristina Fern\'andez de Kirchner} (2003-2015). In 2015, he was elected to integrate the Chamber of Deputies, which finally voted to strip \emph{De Vido} of his congressional immunity over corruption allegations and was immediatly jailed on October 27th.

@@

\par \emph{Milagro Sala} is an indigenous leader. 
She has been incarcerated under pre-trial detention ever since she was first detained in January 2016. She faces allegations of embezzlement related to government funding for housing projects managed by T\'upac Amaru, her social organization.
Sala accused the government of ``violating her human rights", and several people think that she is a political prisoner of the \emph{Mauricio Macri} administration.
